---
title: "Activations cleaning"
author: "Media Innovation Group"
---

## to do

Could directly download [activation data](https://docs.google.com/spreadsheets/d/1vv81AaFwSzLl8lEDYHxI3ANkV2Gf08bh-i_LVDkseH8/edit?usp=sharing). Might require some reworking.

## Overview

It's time to add in a new data source: the heat protocol activation logs. This data was acquired by Lauren McGaughy at KUT through a public information request to the Texas Department of Criminal Justice. 

## Set up

You know the drill!

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(janitor)
library(data.table)
```

## Import activation log

Let's add our activation log file in.

```{r}
raw_activation_log <- read_csv("data-raw/unit-activation-log.csv", skip = 1) |> 
  clean_names()

raw_activation_log
```

## Clean column names

Our column names didn't translate perfectly, so let's work on cleaning them up and remove the last two unnecessary columns.

```{r}
clean_activation_log <- raw_activation_log |> 
  select(
    initial_extreme_temp = initial_date_of_extreme_temperature,
    initiation_date = ics_implementation_date,
    county,
    unit = unit_affected,
    activation_date = date_5, 
    activation_time = time_6,
    deactivation_date = date_7,
    deactivation_time = time_8 
  )
```


## Change date columns

We need our dates to be date columns instead of character columns. Let's adjust.

```{r}
activation_log_dates <- clean_activation_log |> 
  mutate(
    initial_extreme_temp = mdy(initial_extreme_temp),
    initiation_date = mdy(initiation_date),
    activation_date = mdy(activation_date),
    deactivation_date = mdy(deactivation_date)
  ) 

activation_log_dates
```

## Document active days

We need to turn our initial_extreme_temp column, implementation_date column and our deactivation_date column into date ranges.

```{r}
dirty_actives <- activation_log_dates |> 
  group_by(unit) |> 
  mutate(occurence = dense_rank(initial_extreme_temp)) |> 
  mutate(start = as.Date(activation_date), end = as.Date(deactivation_date)) |> 
  mutate(
    activated = map2(start, end, ~seq(from = .x, to = .y, by = "day"))
    ) |> 
  unnest(activated) 

dirty_actives
```

## Clean up our active dates

We need to create a status column, so let's remove everything other than unit and active dates.

```{r}
clean_actives <- dirty_actives |> 
  select(c(unit, activated)) |> 
  mutate(status = "active") |> 
  mutate(date = activated) |> 
  select(!activated)

clean_actives |> 
  filter(unit == "Byrd")
```

## Export

```{r}
clean_actives |> write_rds("data-processed/activation-log.rds")
```




