---
title: "unit-info-cleaning"
---

This data piece will help us join together our unit-focused data and our weather station-focused data. The data was compiled by Media Innovation Group data fellows. The information came from the Texas Department of Criminal Justice website, a TDCJ unit prototype list and the National Weather Service website.

## Goals of this notebook

Clean up our column names so we can easily combine with our other data.

## Loading libraries

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(janitor)
```

## Importing unit info

Let's bring in our unit info sheet. This was compiled by us to be used in this project.

```{r}
units <- read_csv("data-raw/unit-info.csv")

unit_info <- units |> clean_names()

unit_info
```

## Select columns

I'm gonna change column names and only keep the columns we need.

```{r}
unit_info_clean <- unit_info |> 
  rename(unit = unit_name) |> 
  select(unit, type, region, established, station_code)

unit_info_clean
```

## Export data

Our cleaning was pretty easy! Let's export the data and put it into our processed folder now.

```{r}
unit_info_clean |> write_rds("data-processed/unit_info.rds")
```

