---
title: "Indoor temp logs analysis"
author: "Media Innovation Group"
format: html
---

We have recorded indoor temperatures for each unit April through September of 2022, 2023 and 2024 and April through July 25 (day the info request was filled) of 2025. According to the Texas Department of Criminal Justice, indoor temps are taken at 3pm every day from April through September at all units that do not have air conditioning. This information was acquired by Lauren McGaughy at KUT through a public information request to the Texas Department of Criminal Justice.

## Goals of this notebook

In this notebook, we will explore the following questions:

For all units and within individual units:

* How many days was the indoor temperature 85°F or above in 2022? How many consecutive days and when?
* How many days was the indoor temperature 90°F or above in 2022? How many consecutive days and when?
* How many days was the indoor temperature 95°F or above in 2022? How many consecutive days and when?
* How many days was the indoor temperature 100°F or above in 2022? How many consecutive days and when?
* How many days was the indoor temperature 105°F or above in 2022? How many consecutive days and when?

* How many days was the indoor temperature 85°F or above in 2023? How many consecutive days and when?
* How many days was the indoor temperature 90°F or above in 2023? How many consecutive days and when?
* How many days was the indoor temperature 95°F or above in 2023? How many consecutive days and when?
* How many days was the indoor temperature 100°F or above in 2023? How many consecutive days and when?
* How many days was the indoor temperature 105°F or above in 2023? How many consecutive days and when?

* How many days was the indoor temperature 85°F or above in 2024? How many consecutive days and when?
* How many days was the indoor temperature 90°F or above in 2024? How many consecutive days and when?
* How many days was the indoor temperature 95°F or above in 2024? How many consecutive days and when?
* How many days was the indoor temperature 100°F or above in 2024? How many consecutive days and when?
* How many days was the indoor temperature 105°F or above in 2024? How many consecutive days and when?

* How many days was the indoor temperature 85°F or above in 2025? How many consecutive days and when?
* How many days was the indoor temperature 90°F or above in 2025? How many consecutive days and when?
* How many days was the indoor temperature 95°F or above in 2025? How many consecutive days and when?
* How many days was the indoor temperature 100°F or above in 2025? How many consecutive days and when?
* How many days was the indoor temperature 105°F or above in 2025? How many consecutive days and when?

* What month in 2023 reached the highest temperature? What temperature and for how long?
* What month in 2024 reached the highest temperature? What temperature and for how long?
* What month in 2025 reached the highest temperature? What temperature and for how long?

## Setup

Loading our libraries

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(janitor)
library(data.table)
library(ggthemes)
```

## Import

Reading in the combined activation logs and combined indoor logs

```{r}
indoor_logs <- read_rds("data-processed/01-indoor-temps-all-cleaned.rds")
activations_logs <- read_rds("data-processed/01-activation-cleaned.rds")
```

 

## Counting total days of data we have for each year 

```{r}
indoor_logs |> 
  filter(year(date) == 2025) |> 
  group_by(unit) |> 
  summarize(days_recorded = n())
```

Data takeaway: For 2022, 2023 and 2024, we have 183 total days of indoor temperature recordings, which are taken at 3 p.m. each day April 1 through Sept. 30 at all units without air conditioning, according to the TDCJ. As of Oct. 17, 2025, we only have 122 days of 2025 recordings (April 1 through July 25) because we are waiting for the rest of the data to come in via a public information request. 

 
## Counting number of days above each temp per month per unit per year

```{r}
temp_counts_all <- indoor_logs|> 
  group_by(
    unit, flo_mo = floor_date(date, unit = "month")
  ) |> 
  summarize(
    total_days = n(),
    cnt_85 = sum(temperature >= 85, na.rm = T),
    cnt_90 = sum(temperature >= 90),
    cnt_95 = sum(temperature >= 95),
    cnt_100 = sum(temperature >= 100),
    cnt_105 = sum(temperature >= 105)
  )
temp_counts_all

```


## Number of days above each temp per unit per year
 
```{r}
indoor_logs_count <- indoor_logs|> 
  mutate(year = year(date)) |> 
  group_by(unit, year) |> 
  summarize(
    total_days = n(),
    cnt_85 = sum(temperature >= 85, na.rm = T),
    cnt_90 = sum(temperature >= 90, na.rm = T),
    cnt_95 = sum(temperature >= 95, na.rm = T),
    cnt_100 = sum(temperature >= 100, na.rm = T)
  ) 

indoor_logs_count |> filter(year == 2022)
```
 
## Plotting days above each temp per unit per year

Note: there is a substantial dip in the number of days recorded in 2025 because the dataset is only complete through July 25. 

85° and above

```{r}
ggplot(
  indoor_logs_count,
aes(x = year, y= cnt_85, group = unit)
) +
geom_line(aes(group = unit)) +
  theme(legend.position = "none") +
  labs(
    title = "Number of days with indoor temperature recordings at or above 85°F",
    subtitle = str_wrap("According to the Texas Department of Criminal Justice, indoor temperature recordings are taken at 3 p.m. every day from April 1 to Sept. 30 at all units that do not have air conditioning.There are 183 days of temperature recordings in 2022, 2023 and 2024, and 122 days of recordings in 2025."),
    x = "Year", y = "Number of days at or above 85°F",
    caption = "Source: Texas Department of Criminal Justice"
  ) +
  theme_bw()

```

90° and above

```{r}
ggplot(
  indoor_logs_count,
aes(x = year, y= cnt_90, group = unit)
) +
geom_line(aes(group = unit)) +
  theme(legend.position = "none") +
  labs(
    title = "Number of days with indoor temperature recordings at or above 90°F",
    subtitle = str_wrap("According to the Texas Department of Criminal Justice, indoor temperature recordings are taken at 3 p.m. every day from April 1 to Sept. 30 at all units that do not have air conditioning. There are 183 days of temperature recordings in 2022, 2023 and 2024, and 122 days of recordings in 2025."),
    x = "Year", y = "Number of days at or above 90°F",
    caption = "Source: Texas Department of Criminal Justice"
  ) +
  theme_bw()
```

95° and above

```{r}
ggplot(
  indoor_logs_count,
aes(x = year, y= cnt_95, group = unit)
) +
geom_line(aes(group = unit)) +
  theme(legend.position = "none") +
  labs(
    title = "Number of days with indoor temperature recordings at or above 95°F",
    subtitle = str_wrap("According to the Texas Department of Criminal Justice, indoor temperature recordings are taken at 3 p.m. every day from April 1 to Sept. 30 at all units that do not have air conditioning. There are 183 days of temperature recordings in 2022, 2023 and 2024, and 122 days of recordings in 2025."),
    x = "Year", y = "Number of days at or above 95°F",
    caption = "Source: Texas Department of Criminal Justice"
  ) +
  theme_bw()
```

100° and above


```{r}
ggplot(
  indoor_logs_count,
aes(x = year, y= cnt_100, group = unit)
) +
geom_line(aes(group = unit)) +
  theme(legend.position = "none") +
  labs(
    title = "Number of days with indoor temperature recordings at or above 100°F",
    subtitle = str_wrap("According to the Texas Department of Criminal Justice, indoor temperature recordings are taken at 3 p.m. every day from April 1 to Sept. 30 at all units that do not have air conditioning. There are 183 days of temperature recordings in 2022, 2023 and 2024, and 122 days of recordings in 2025."),
    x = "Year", y = "Number of days at or above 100°F",
    caption = "Source: Texas Department of Criminal Justice"
  ) +
  theme_bw()
```
```


 
## 2022 indoor logs analysis
 
## Days at or above 85°

Filtering the indoor log data to include only readings taken in 2022 where the temperature was above 85°F, grouping by unit and summarizing by number of appearances.

```{r}
indoor_2022_85 <- indoor_logs |> 
  filter(
    temperature >= 85
    ) |> group_by(unit, year(date)) |> 
  filter(year(date) == 2022) |> 
  summarize(days_above_85 = n()) |> 
  arrange(desc(days_above_85)) 

indoor_2022_85
```

Data takeaway: In 2022, 67 out of 68 units without air conditioning recorded indoor temperatures at or above 85 degrees for at least 43 days between April 1 and Sept. 30. The Sayle unit, located in Breckenridge, Texas, was the only unit to not record temperatures at or above 85 degrees. 51 units recorded indoor temperatures at or above 85 degrees for over 50% of days, or at least 92 days. Ten units recorded indoor temperatures at or above 85 degrees for over 75% of days, or at least 138 days. The Garza West unit, located in Beeville, Texas, recorded the most days with indoor temperatures at or above 85 degrees, with 168 days, or 92% of days between April 1 and Sept. 30.  

## Days at or above 90

Filtering the indoor log data to include only readings taken in 2022 where the temperature was above 90°F, grouping by unit and summarizing by number of appearances.

```{r}
indoor_2022_90 <- indoor_logs |> 
  filter(
    temperature >= 90
    ) |> group_by(unit, year(date)) |> 
  filter(year(date) == 2022) |> 
  summarize(days_above_90 = n()) |> 
  arrange(desc(days_above_90)) 

indoor_2022_90
```

Data takeaway: In 2022, 65 out of 68 units without air conditioning recorded indoor temperatures at or above 90 degrees at least once between April 1 and Sept. 30. 41 units recorded indoor temperatures at or above 90 degrees for over 25% of days, or at least 46 days. Nine units recorded indoor temperatures at or above 90 degrees for over 50% of days, or at least 92 days. The Dominguez unit, located in San Antonio, Texas, recorded the most days with indoor temperatures at or above 90 degrees, with 122 days, or two-thirds of days between April 1. and Sept. 30. 

## Days at or above 95

Filtering the indoor log data to include only readings taken in 2022 where the temperature was above 95°F, grouping by unit and summarizing by number of appearances.

```{r}
indoor_2022_95 <- indoor_logs |> 
  filter(
    temperature >= 95
    ) |> group_by(unit, year(date)) |> 
  filter(year(date) == 2022) |> 
  summarize(days_above_95 = n()) |> 
  arrange(desc(days_above_95))

indoor_2022_95
```

Data takeaway: In 2022, 51 out of 68 units without air conditioning recorded indoor temperatures at or above 95 degrees at least once between April 1 and Sept. 30. Six units recorded indoor temperatures at or above 95 degrees for over 25% of days, or at least 46 days. The Stevenson unit, located in Cuero, Texas, recorded the most days with indoor temperatures at or above 95 with 74 days, or over 40% of days between April 1. and Sept. 30. 

## Days at or above 100

Filtering the indoor log data to include only readings taken in 2022 where the temperature was above 100°F, grouping by unit and summarizing by number of appearances.


```{r}
indoor_2022_100 <- indoor_logs |> 
  filter(
    temperature >= 100
    ) |> group_by(unit, year(date)) |> 
  filter(year(date) == 2022) |> 
  summarize(days_above_100 = n()) |> 
  arrange(desc(days_above_100)) 

indoor_2022_100

indoor_logs |> filter(
  unit == "Stevenson",
  temperature >= 100
)
```

Data takeaway: In 2022, 15 out of 68 units without air conditioning recorded indoor temperatures at or above 100 degrees at least once between April 1 and Sept. 30. The Stevenson unit, located in Cuero, Texas, recorded the most days with indoor temperatures at or above 100 degrees with 13 days between April 1 and Sept. 30. This includes four consecutive days between July 10 and July 13. 

## 2023 indoor logs analysis

## Days at or above 85°

Filtering the indoor log data to include only readings taken in 2023 where the temperature was above 85°F, grouping by unit and summarizing by number of appearances.

```{r}
indoor_2023_85 <- indoor_logs |> 
  filter(
    temperature >= 85
    ) |> group_by(unit, year(date)) |> 
  filter(year(date) == 2023) |> 
  summarize(days_above_85 = n()) |> 
  arrange(desc(days_above_85)) 

indoor_2023_85
```

