---
title: "Activations cleaning"
author: "Media Innovation Group"
---

## Overview

These are the days that a unit was under ICS protocols like providing extra ice water and access to cooler areas during extreme heat events. This data was acquired by Lauren McGaughy of The Texas Newsroom through a public information request to the Texas Department of Criminal Justice.

## Possible update

In our tracking Google sheet we do have the same data, except there are errors on the last line.

## Set up

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(janitor)
library(data.table)
```

## Comparison activation log

We'll check about the original CSV with Lauren, but in the meantime we'll compare that with the Activation records we also have in our tracking sheet. This copy was downloaded from our [MIG = TDCJ Basic Information](https://docs.google.com/spreadsheets/d/1vv81AaFwSzLl8lEDYHxI3ANkV2Gf08bh-i_LVDkseH8/edit?gid=1968960884#gid=1968960884) on the "Activation" tab. The Google Sheet has errors on the last record for 9/5/2023, but that record is in full in our csv log.

```{r}
#| label: download-activation-direct
#| eval: false

# download.file(
#   "https://docs.google.com/spreadsheets/d/e/2PACX-1vRebFs9O2i0HoygXTtIvuzdVTmyrjX3MeUorU9d4fpYkGX7Bb026OradFQ1MMk2ltcGnyILih6ow4F4/pub?gid=1968960884&single=true&output=csv",
#   "data-original/unit-activation-direct.csv"
#       )
```

Import direct log

```{r}
raw_activation_direct <- read_csv("data-original/unit-activation-direct.csv") |> 
  clean_names()

raw_activation_direct
```

This original files lacks on activation that is in our already-downloaded `unit-activation-log.csv`. We'll continue to use the -log version while we track down an original.

## Import activation log

Let's add our activation log file in. (This file was downloaded directly into the repo, but we lost track of the original file. It has an additional record or other copies don't have.)

```{r}
raw_activation_log <- read_csv("data-original/unit-activation-log.csv", skip = 1) |> 
  clean_names()

raw_activation_log
```

## Clean column names

Our column names didn't translate perfectly, so let's work on cleaning them up and remove the last two unnecessary columns.

```{r}
clean_activation_log <- raw_activation_log |> 
  select(
    initial_extreme_temp = initial_date_of_extreme_temperature,
    initiation_date = ics_implementation_date,
    county,
    unit = unit_affected,
    activation_date = date_5, 
    activation_time = time_6,
    deactivation_date = date_7,
    deactivation_time = time_8 
  )

```


## Change date columns

We need our dates to be date columns instead of character columns. Let's adjust.

```{r}
activation_log_dates <- clean_activation_log |> 
  mutate(
    initial_extreme_temp = mdy(initial_extreme_temp),
    initiation_date = mdy(initiation_date),
    activation_date = mdy(activation_date),
    deactivation_date = mdy(deactivation_date)
  ) 

activation_log_dates
```

## Document active days

We need to turn our initial_extreme_temp column, implementation_date column and our deactivation_date column into date ranges.

```{r}
dirty_actives <- activation_log_dates |> 
  group_by(unit) |> 
  mutate(occurence = dense_rank(initial_extreme_temp)) |> 
  mutate(start = as.Date(activation_date), end = as.Date(deactivation_date)) |> 
  mutate(
    activated = map2(start, end, ~seq(from = .x, to = .y, by = "day"))
    ) |> 
  unnest(activated) 

dirty_actives
```

## Clean up our active dates

We need to create a status column, so let's remove everything other than unit and active dates.

```{r}
clean_actives <- dirty_actives |> 
  select(c(unit, activated)) |> 
  mutate(protocol_active = T) |> 
  mutate(date = activated) |> 
  select(!activated)

clean_actives |> 
  filter(unit == "Byrd")
```

## Export

```{r}
clean_actives |> write_rds("data-processed/01-activation-cleaned-old.rds")
```




